---
title: "CUB SDM Report v2.5"
author: "Avery Hill"
date: "`r Sys.Date()`"
output: 
  rmdformats::robobook:
    # toc: true
    # toc_float: true
    toc_depth: 6
runtime: shiny
resource_files:
  - data
  
---

<style>
.html-widget {
margin: auto;
}

.label-left .form-group {
display: flex;              /* Use flexbox for positioning children */
flex-direction: row;        /* Place children on a row (default) */
width: 100%;                /* Set width for container */
max-width: 400px;
}

.label-left label {
margin-right: 2rem;         /* Add spacing between label and slider */
align-self: center;         /* Vertical align in center of row */
text-align: right;
flex-basis: 100px;          /* Target width for label */
}

.label-left .irs {
flex-basis: 300px;          /* Target width for slider */
}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = F,
  echo = F,
  warning = F,
  fig.align = 'center'
)
```

```{r load-packages, include=F}
library(shiny)
library(leaflet)
library(sf)
library(dplyr)
library(terra)
# library(gt)
library(duckdb)

```


```{r, include=F}
# Load in duckdb and everything
```

# Mount Tam

