---
title: "New Phytologist Areas of Interest"
# author: "Avery Hill"
date: "`r Sys.Date()`"
# output: 
  # rmdformats::robobook:
    # toc: true
    # toc_float: true
    # toc_depth: 6
runtime: shiny
resource_files:
  - data
  
---

<style>
.html-widget {
margin: auto;
}

.label-left .form-group {
display: flex;              /* Use flexbox for positioning children */
flex-direction: row;        /* Place children on a row (default) */
width: 100%;                /* Set width for container */
max-width: 400px;
}

.label-left label {
margin-right: 2rem;         /* Add spacing between label and slider */
align-self: center;         /* Vertical align in center of row */
text-align: right;
flex-basis: 100px;          /* Target width for label */
}

.label-left .irs {
flex-basis: 300px;          /* Target width for slider */
}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = F,
  echo = F,
  warning = F,
  fig.align = 'center'
)
```

```{r load-packages, include=F}
library(shiny)
library(leaflet)
library(sf)
library(dplyr)
library(terra)
# library(gt)
library(duckdb)
library(dbplyr)
library(tidyr)
# library(duckdbfs)
library(readr)
library(purrr)
library(shinyWidgets)
library(DT)
library(stringr)

options(scipen = 999)
```


```{r, include=F}
# Load in duckdb and occurrences
dcon <- dbConnect(duckdb())
dcon |> dbExecute("INSTALL spatial; LOAD spatial;")
dcon |> dbExecute("
CREATE VIEW occurrences AS
SELECT * EXCLUDE geom, ST_Point(decimallongitude, decimallatitude) AS geom
FROM read_parquet('occurrences.parquet');")

# Places
# dcon |> dbExecute("
# CREATE VIEW places AS
# SELECT * EXCLUDE geom, ST_ASTEXT(ST_GEOMFROMWKB(geom)) AS geom
# FROM read_parquet('place_boundaries.parquet');")
dcon |> dbExecute("
CREATE VIEW places AS
SELECT *
FROM read_parquet('place_boundaries.parquet');")

```

---

```{r}
# Place selection
places <- dcon |> 
  tbl("occurrences") |> 
  distinct(place_name) |> 
  pull(1)

fluidRow(
  radioGroupButtons(
   inputId = "place",
   # label ="Deduplicate?",
   selected = "Marble/Salmon Mountains",
   width = "75vw",
   choices = places,
   justified = TRUE
)
  )
```


```{r}
fluidRow(
  "Note: udpated 2024-11-25, fixed duplicates and recordedby column",
  h1(renderText(input$place)),
  wellPanel(
    h5("Remove duplicate records?"),
    radioGroupButtons(
     inputId = "dup_choice",
     # label ="Remove duplicate records?",
     choices = c("yes", "no"),
     selected = "no"
     # justified = TRUE
     ),
    h5("Observation type"),
    radioGroupButtons(
     inputId = "obs_type",
     choices = c("All", 
                 "iNaturalist" = "HUMAN_OBSERVATION", 
                 "Herbaria"  =  "PRESERVED_SPECIMEN"),
     selected = "All",
     width = "50vw",
     justified = TRUE
     ),
    h5("Filter by coordinate uncertainty?"),
    radioGroupButtons(
     inputId = "uncert_filter",
     choices = c("No Filter", 
                 "Less than 500m",
                 "Less than 1000m",
                 "Less than 25km"),
     selected = "No Filter",
     width = "50vw",
     justified = TRUE
     )
    )
  )
```

```{r}
x <- dcon |> tbl("occurrences") |>
  colnames()
  # filter(institutioncode == "iNaturalist")

# dcon |> tbl("occurrences") |> select(eventdate) |> arrange(desc(eventdate))

# dcon |> tbl("occurrences") |>
#   # filter(institutioncode == "iNaturalist") |>
#   select(species, verbatimscientificname) |> 
#   distinct() |> 
#   collect() |> 
#   # This takes care of taxa with ' × ' between genus and species
#   mutate(verbatim_as_binomial = str_replace_all(verbatimscientificname, "\\s+\\w\\s+", " ")) |> 
#   # mutate(verbatim_as_binomial = sql("REGEXP_EXTRACT(verbatimscientificname, '^\\w+\\s+\\w+(?:-\\w+)?')")) |> 
#     mutate(verbatim_as_binomial = str_extract(verbatimscientificname, "^\\w+\\s+\\w+(?:-\\w+)?")) |> 
#   rename(gbif_binomial = species) |> 
#   filter(gbif_binomial != verbatim_as_binomial) |> 
#   select(
#     gbif_binomial,
#          verbatim_as_binomial,
#     verbatimscientificname
#          ) |> 
#   distinct(gbif_binomial, verbatim_as_binomial, .keep_all = T) |> 
#         arrange(gbif_binomial) |> 
#   # collect() |> 
#     View()
# 

```


```{r}
mar_occ <- reactive({
  
  req(input$place, input$dup_choice, input$obs_type)

  occ <- dcon |> 
    tbl("occurrences") |> 
    filter(place_name == input$place,
           basisofrecord %in% c("HUMAN_OBSERVATION", "PRESERVED_SPECIMEN")) |> 
    mutate(gbif_record_link = paste0("https://www.gbif.org/occurrence/", gbifid),
    recordedby = as.character(recordedby))
  
  if(input$dup_choice == "yes"){
    
    occ <- occ |> 
      # distinct(verbatimscientificname, 
      # decimallatitude, decimallongitude, 
      # year, month, day, 
      #          collectioncode, 
      #          institutioncode, .keep_all = T)
distinct( 
      decimallatitude, decimallongitude, basisofrecord,
      year, month, day, species, recordedby, .keep_all = T)
      # distinct(species, decimallatitude, decimallongitude, year, month, day,
      #           place_name, stateprovince, coordinateuncertaintyinmeters,
      #           basisofrecord, geom, .keep_all = T) # notice that institutioncode is out
    
  }
  
  # print(input$obs_type)
  if(input$obs_type != "All"){
    chosen_obs_type <- input$obs_type
    occ <- occ |> 
      filter(basisofrecord == chosen_obs_type)
  }
  
  # Apply the filter for uncertainty if it is not "All"
  if (input$uncert_filter == "No Filter") {
    # Do nothing, no further filtering
  } else if (input$uncert_filter == "Less than 500m") {
    occ <- occ |>
      filter(coordinateuncertaintyinmeters <= 500 | is.na(coordinateuncertaintyinmeters))
  } else if (input$uncert_filter == "Less than 1000m") {
    occ <- occ |>
      filter(coordinateuncertaintyinmeters <= 1000 | is.na(coordinateuncertaintyinmeters))
  } else if (input$uncert_filter == "Less than 25km") {
    occ <- occ |>
      filter(coordinateuncertaintyinmeters <= 25000 | is.na(coordinateuncertaintyinmeters))
  }
  
  # print(occ)
  
  return(occ)
})

  

herb_n_inat_count <- reactive({
  req(mar_occ())
  mar_occ() |>
  group_by(basisofrecord) |> summarise(N = n()) |>
  # filter(basisofrecord %in% c("HUMAN_OBSERVATION", "PRESERVED_SPECIMEN")) |>
  pivot_wider(names_from = basisofrecord, values_from = N) |>
  collect()
})

herb_n_inat_speccount <- reactive({
  req(mar_occ())
  mar_occ() |>
  group_by(basisofrecord) |> 
    distinct(species) |> 
    summarise(N = n()) |>
  # filter(basisofrecord %in% c("HUMAN_OBSERVATION", "PRESERVED_SPECIMEN")) |>
  pivot_wider(names_from = basisofrecord, values_from = N) |>
  collect()
})

herb_n_inat_totalspecies <- reactive({
  req(mar_occ())
  # browser()
  mar_occ() |>
    # filter(basisofrecord %in% c("HUMAN_OBSERVATION", "PRESERVED_SPECIMEN")) |> 
    distinct(species) |>
    count() |> 
    collect()
})
  

```

#### Observations
There are `r renderText(herb_n_inat_count()$PRESERVED_SPECIMEN + herb_n_inat_count()$HUMAN_OBSERVATION)` total observations,
`r renderText((100 * herb_n_inat_count()$PRESERVED_SPECIMEN / (herb_n_inat_count()$PRESERVED_SPECIMEN + herb_n_inat_count()$HUMAN_OBSERVATION)) |> round(1))`% of which are herbarium records.

(`r renderText(herb_n_inat_count()$PRESERVED_SPECIMEN)` herbarium records; `r renderText(herb_n_inat_count()$HUMAN_OBSERVATION)` iNaturalist records)

#### Species
There are `r renderText(herb_n_inat_totalspecies()$n)` total species,
`r renderText((100 * herb_n_inat_speccount()$PRESERVED_SPECIMEN / (herb_n_inat_totalspecies()$n)) |> round(1))`% of which are found in herbarium records.

(`r renderText(herb_n_inat_speccount()$PRESERVED_SPECIMEN)` species in herbarium records; `r renderText(herb_n_inat_speccount()$HUMAN_OBSERVATION)` species in iNaturalist records)




```{r}

output$map <- renderLeaflet({
  rast_occ <- mar_occ() |>
    mutate(
      latitude = round(decimallatitude, 2),
      longitude = round(decimallongitude, 2)
    ) |>
    count(longitude, latitude
          # basisofrecord,
          # institutioncode, 
          #coordinateuncertaintyinmeters
          ) |>
  collect()

count_rast <- rast_occ |> rast(crs = "epsg:4326")

boundary <- dcon |> 
  tbl("places") |> 
  filter(name == input$place) |>
  collect() |> 
  st_as_sf(wkt = "geom")

res_km <- res(count_rast) * 111

pal <- colorNumeric("viridis",
                    values(count_rast[["n"]]),
                    na.color = "transparent")
leaflet() |>
  addProviderTiles("CartoDB.Positron") |>
  addRasterImage(count_rast[["n"]],
                 colors = pal) |>
  addPolygons(data = boundary,
              color = "blue",  # Set polygon border color
              fillOpacity = 0,
              weight = 2) |>
  addLegend(
    pal = pal,
    values = values(count_rast[["n"]]),
    title = sprintf("Obs. Count / %skm", res_km[1] |> round(2)),
    position = "bottomright"
  )
})

leafletOutput("map")

```

#### Download currently filtered data?
```{r}
# Download
fluidRow(downloadButton("download_gpkg", "Download as GPKG"),
      downloadButton("download_csv", "Download as CSV"),
downloadButton("download_spec", "Download species lists"))

 output$download_gpkg <- downloadHandler(
    filename = function() {
      paste0(input$place, "_occurrences", ".gpkg")
    },
    content = function(file) {
      # Write the sf to GPKG format
      clean_occ <- mar_occ() |>
        select(-geom) |>
        collect() |>
        # Need to clean lists out
        mutate(across(where(is.list), ~ map(., toString))) |>
        unnest(cols = where(is.list)) |>
        st_as_sf(coords = c("decimallongitude", "decimallatitude"),
           crs = 4326)
      write_sf(clean_occ, dsn = file)
    }
  )

  # CSV Download
  output$download_csv <- downloadHandler(
    filename = function() {
      paste0(input$place, "_occurrences", ".csv")
    },
    content = function(file) {
      # Write the tibble to CSV format

      write_csv(mar_occ() |> select(-geom) |> collect(), file)
    }
  )
  
  # Species lists download
  output$download_spec <- downloadHandler(
    filename = function() {
      paste0(input$place, "_species_list", ".csv")
    },
    content = function(file) {
      # Write the tibble to CSV format

      write_csv(mar_occ() |> 
                  distinct(species, basisofrecord) |> 
                  arrange(species) |> 
                  collect(), 
                file)
    }
  )

```

# Taxon name changes
I did some `regex` magic to try and help identify which species names were changed to conform to the GBIF backbone taxonomy. 
I essentially did this by gluing the columns `species` and `infraspecificepithet` together into `modified_gbif_backbone`, and then comparing this to the `verbatimscientificname` with the author names removed (now called `modified_verbatim_name`).
Then I filtered for all rows where `modified_verbatim_name` and `verbatimscientificname` did not match.
I included the constituent columns to the right so you can see what the original taxon information was.

It's not perfect but it might be helpful. You can download the table below.

At a rough approximation in this filtered dataset, **`r renderText(verbatim_change_taxon_count())`** verbatim taxon names were standardized to **`r renderText(gbif_change_taxon_count())`** gbif backbon taxon names.
```{r}

fluidRow(
  radioGroupButtons(
     inputId = "namechange_choice",
     # label ="Remove duplicate records?",
     choices = c("Binomial", "Trinomial"),
     selected = "Binomial"
     # justified = TRUE
     )
)

# Species table
namechange_table <- reactive({
  req(mar_occ())
  
  if(input$namechange_choice == "Trinomial"){
    mar_occ() |> 
      mutate(modified_gbif_backbone = paste(species, infraspecificepithet),
         modified_verbatim_name = verbatimscientificname |> 
           # Get rid of the modifiers
           str_replace_all("\\s*(subsp\\.|var\\.|\\sex\\s|\\sx\\s|&|\\s×\\s|\\sbis\\s|\\svar\\s|\\sf\\.\\s)\\s*", " ") |> # need to work on this a little more
           # Remove everything in parantheses
           str_replace_all("\\s*\\([^)]*\\)", "") |>
           # Get rid of words (names) that are capitalized after the first one
           str_replace_all(" [A-Z][^ ]*", "") |>
           str_squish()
         ) |> 
  filter(modified_gbif_backbone != modified_verbatim_name) |> 
  select(
    modified_gbif_backbone,
         modified_verbatim_name,
    species,
         infraspecificepithet, 
    # scientificname,
            verbatimscientificname
         ) |> 
  distinct(modified_gbif_backbone, modified_verbatim_name, .keep_all = T) |> 
        arrange(modified_gbif_backbone) |> 
  collect()
    
  } else if (input$namechange_choice == "Binomial") {
    
    mar_occ() |> 
      select(species, verbatimscientificname) |> 
  distinct() |> 
  collect() |> 
  # This takes care of taxa with ' × ' between genus and species
  mutate(verbatim_as_binomial = str_replace_all(verbatimscientificname, 
                                                "\\s+\\w\\s+", " ")) |> 
    mutate(verbatim_as_binomial = str_extract(verbatimscientificname, "^\\w+\\s+\\w+(?:-\\w+)?")) |> 
  rename(gbif_binomial = species) |> 
  filter(gbif_binomial != verbatim_as_binomial) |> 
  select(
    gbif_binomial,
         verbatim_as_binomial,
    verbatimscientificname
         ) |> 
  distinct(gbif_binomial, verbatim_as_binomial, .keep_all = T) |> 
        arrange(gbif_binomial)
  }
})

# Get number of taxa that changed names
gbif_change_taxon_count <- reactive({
  if(input$namechange_choice == "Trinomial"){
    namechange_table() |> distinct(modified_gbif_backbone) |> nrow()
  } else if (input$namechange_choice == "Binomial") {
    namechange_table() |> distinct(gbif_binomial) |> nrow()
  }
})

verbatim_change_taxon_count <- reactive({
    if(input$namechange_choice == "Trinomial"){
    namechange_table() |> distinct(modified_verbatim_name) |> nrow()
  } else if (input$namechange_choice == "Binomial") {
    namechange_table() |> distinct(verbatim_as_binomial) |> nrow()
  }
})

output$species_table <- renderDT({
    
    datatable(namechange_table(), 
              options = list(scrollY = "300px", paging = FALSE))  # Scrollable table
  })

DTOutput("species_table") 

fluidRow(downloadButton("download_namechange", "Download name change table (csv)"))

# Now for download
output$download_namechange <- downloadHandler(
    filename = function() {
      paste0(input$place, "_namechange", ".csv")
    },
    content = function(file) {
      # Write the tibble to CSV format

      write_csv(namechange_table(), 
                file)
    }
  )
```

---
##### Notes
The column 'basisofrecord' is the GBIF column that indicates how the data were collected.
There are two values here, HUMAN_OBSERVATION and PRESERVED_SPECIMEN.
These practically correspond to iNaturalist and herbaria, respectively.

These are the filters I used to pre-filter the data (in SQL):
```sql
SELECT *
  FROM gbif
  WHERE phylum = 'Tracheophyta'
  AND species IS NOT NULL
  AND NOT species = ''
  AND stateprovince = 'California'
```
- These data were downloaded from GBIF on August 20th, 2024
